Optimization of String Rewriting Operations for 3D Fractal Generation with Genetic Algorithms

### 2018-07-28 Optimization-of-String-Rewriting-Operations-for-3D-Fractal-Generation-with-Genetic-Algorithms/paper.tex
Final fixes were done.
Small mistake was fixed.
Initial version of the presentation was done.

This study addresses IFP which is well known in the fractals theory \cite{guerin01,nettleton01}. There is a branch in the theory for the formal languages which is related to a context-free grammar (CFG). In such grammar there are set of production rules. Application of these rules, lead to generation of all possible strings in a specified formal language \cite{ochoa01}. In CFG, a production rule is an operation of simple replacement. Rules are applied regardless of the context. The left hand side of a CFG rule is always nonterminal symbol. It means that nonterminal symbols are not part of the resulting string. The right hand side of a CFG rule, consists of terminal symbol, nonterminal symbols or combination of both. The main idea in CFGs is that nonterminal symbols are substituted with the given rules until there is no nonterminal symbols in the generated string.

@@

The problem in 1D is well presented by \cite{shonkwiler01}, but in Hodkinson's research, the genetic algorithms (GAs) are used for 2D shape of Greek letter $\pi$, to be reconstructed with substitution rules of 30 shades of blue pixels (Fig. \ref{fig:pic01}). The usage of differential evolution is also applicable \cite{zelinka01}, but not in Hodkinson's research, because the optimization is in a discrete space.  The key problem in this research is which rules to be selected to achieved the final goal. GAs is a promising approach for such kind of combinatorial optimization. A single rule has LHS pixel with particular color and nine pixels on the RHS as 3x3 substitution matrix. Image starts with a single pixel and it grows on size by 3 on each iteration.

@@

The optimization process converges much more slower in the 3D space than what was achieved by Hodkinson in the 2D space. As it is shown in Fig. \ref{fig:pic03}, starting form a single cube it goes to the desired shape too slowly.

>>>>

This study addresses FIP which is well known in the fractals theory \cite{guerin01,nettleton01}. There is a branch in the theory for the formal languages which is related to a context-free grammar (CFG). In such grammar there are set of production rules. Application of these rules, lead to generation of all possible strings in a specified formal language \cite{ochoa01}. In CFG, a production rule is an operation of simple replacement. Rules are applied regardless of the context. The left hand side of a CFG rule is always nonterminal symbol. It means that nonterminal symbols are not part of the resulting string. The right hand side of a CFG rule, consists of terminal symbol, nonterminal symbols or combination of both. The main idea in CFGs is that nonterminal symbols are substituted with the given rules until there is no nonterminal symbols in the generated string.

@@

The problem in 1D is well presented by \cite{shonkwiler01}, but in Hodkinson's research, the genetic algorithms (GAs) are used for 2D shape of Greek letter $\pi$, to be reconstructed with substitution rules of 30 shades of blue pixels (Fig. \ref{fig:pic01}). The usage of differential evolution is also applicable \cite{zelinka01}, but not in Hodkinson's research, because the optimization is in a discrete space.  The key problem in this research is which rules to be selected to achieved the final goal. GA is a promising approach for such kind of combinatorial optimization. A single rule has LHS pixel with particular color and nine pixels on the RHS as 3x3 substitution matrix. Image starts with a single pixel and it grows on size by 3 on each iteration.

@@

The optimization process converges much more slower in the 3D space than what was achieved by Hodkinson in the 2D space. As it is shown in Fig. \ref{fig:pic03}, starting form a single cube it goes to the desired shape too slowly. The target shape is sphere, but intermediate shapes looks more as cubes than the desired target object.

### 2018-07-28 Optimization-of-String-Rewriting-Operations-for-3D-Fractal-Generation-with-Genetic-Algorithms/presentation.tex
Initial version of the presentation was done.

@@

\subsection{Advantages and Disadvantages}

\subsection{Q\&A}

>>>>

\begin{frame}
\center \huge{Introduction}
\end{frame}

\begin{frame}
\frametitle{Fractal Inverse Problem}
\begin{itemize}
  \item It is an important research area with a great number of potential application fields
  \item It consists in finding a fractal model or code that generates a given object
\end{itemize}
Definition by Eric Guerin and Eric Tosan
\end{frame}

\begin{frame}
\frametitle{String Rewriting System}
\begin{itemize}
  \item A substitution system in which rules are used to operate on a string consisting of letters of a certain alphabet
  \item String rewriting is a particularly useful technique for generating successive iterations of certain types of fractals
\end{itemize}
Definition by Wolfram MathWorld
\end{frame}

\begin{frame}
\frametitle{2D Fractal by Jack Hodkinson}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.65\linewidth]{pic01}
\label{figure01}
\end{figure}
\end{frame}

\section{Model and Optimization}

\begin{frame}
\center \huge{Model and Optimization}
\end{frame}

\begin{frame}
\frametitle{3D Fractal Generation}
\begin{itemize}
  \item A finite 3D space as three-dimensional array of voxels
  \item Each voxel has RGB value (red, green, blue) for the color
  \item 32 transition matrices are used (3x3x3 as dimensions)
  \item 32 colors shades of gray are used
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Fractal Generation Algorithm}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.95\linewidth]{pic04}
\label{figure02}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Generation Algorithm}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.95\linewidth]{pic05}
\label{figure03}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Optimization Specifics}
\begin{itemize}
  \item GA individuals are sets of 32 transition matrices
  \item Parents for the crossover are appointed by tournament selection with arity of two
  \item Uniform crossover is used
  \item Random change of a color is used as mutation
  \item Euclidean distance between generated fractal and target shape is used as fitness function
  \item Time limit is used as stopping criteria
\end{itemize}
\end{frame}

\section{Experiments and Results}

\begin{frame}
\center \huge{Experiments and Results}
\end{frame}

\begin{frame}
\frametitle{Experimental Setup}
\begin{itemize}
  \item All experiments are done in Java as a programming language
  \item Apache Genetic Algorithms Framework is used for the optimization
  \item Printing in 3D - JavaSCAD is used as 3D visualization tool
  \item The source code is available at: github.com/TodorBalabanov/Jack-Hodkinson-3D-Fractal
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Genetic Algorithm Parameters}
\begin{table}[h!]
\centering
\label{table01}
\begin{tabular*}{\textwidth}{|c@{\extracolsep{\fill}}|c|}
\hline 
\textbf{Parameter} & \textbf{Value} \\
\hline
\hline
elitism rate & 0.10 \\
\hline
crossover rate & 0.90 \\
\hline
mutation rate & 0.01 \\
\hline
number of individuals & 37 \\
\hline
number of variables &  864 \\
\hline
optimization minutes & 30 \\
\hline
\end{tabular*}
\end{table}
\end{frame}

\begin{frame}
\frametitle{Target Object}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.75\linewidth]{pic02}
\label{figure04}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Intermediate Objects}
\begin{figure}[h]
  \centering
  \includegraphics[width=1.00\linewidth]{pic03}
\label{figure05}
\end{figure}
\end{frame}

@@

\begin{frame}
\center \huge{Conclusions}
\end{frame}

\begin{frame}
\frametitle{Concluding Remarks}
\begin{itemize}
  \item Such optimization is a time consuming process
  \item Virtual 3D objects could be better presented as real world objects with 3D scanner and 3D color printer
  \item A possible application of the presented ideas is compression of 3D images
  \item As further research, parallelization of the genetic algorithm can lead to optimization speed-up
\end{itemize}
\end{frame}
